resistanceModelNotesAndy

### 2015-05-01 resistanceModelNotesAndy.Rmd
progress notes

**30/4/15** 9.45-

Need to be careful about different W fitness variables
`W.RR1_00 : [locus][niche]` single locus (Wlocus)
`W.RR1SS2_0b :   [locus1][locus2][niche]` niche (Wniche)
`W.m.SS1SS2 : [sex][locus1][locus2]` sex (Windiv)

@@

@@

@@

Different W fitness variables
`W.RR1_00 : [locus][niche]` single locus
`W.RR1SS2_0b :   [locus1][locus2][niche]` niche
`W.m.SS1SS2 : [sex][locus1][locus2]` sex

>>>>

**30/4/15** 9.45-13.30 3.75hrs 2.15-5.30 3.25hrs 6-7 8hrs

Need to be careful about different W fitness variables
`W.RR1_00 : [locus,niche]` single locus (Wlocus)
`W.RR1SS2_0b :   [locus1, locus2, niche]` niche (Wniche)
`W.m.SS1SS2 : [sex, locus1, locus2]` sex (Windiv)

@@

So now I can keep developing runModel2() and use Ctrl T to test I'm not breaking it.

Think, does my Wniche contain too many elements at 81 ?
i.e. does it contain impossible combinations ? I think I'm OK Beths code fills 9*9 groups of individual variables.

`W.RR1SS2_0b` ... `*81 3*3*3*3` to: Wniche[locus1, locus2, niche1, niche2]

I may need to add a genotype option to createArray() to cope with f that needs cis & trans

If I convert niche_0B to niche[niche1,niche2] I can simplify further the calculation of 
Two genotype fitnesses in two insecticide Niche
by having a loop that goes through each niche.

Coolio! reduced 250 lines of code to ~ 20 with this
```
    ## Two genotype fitnesses in two insecticide Niche ##

    #!r to replace 250+ lines below
    for( niche1 in dimnames(Wniche)$niche1)
    {
      for( niche2 in dimnames(Wniche)$niche2)
      {
        #if this niche toggled off set fitness to 0
        if (niche[niche1,niche2] == 0)
        {
          Wniche[,,niche1,niche2] <- 0
        } else{
          #otherwise set fitness to product of the 2 loci
          for( locus1 in dimnames(Wniche)$locus1)
          {
            for( locus2 in dimnames(Wniche)$locus2)
            {    
              Wniche[locus1,locus2,niche1,niche2] <- Wloci[locus1,niche1,niche2] * Wloci[locus2,niche1,niche2]
            }
          }          
        }
      }
    }
```

Trying to get fitnesses from my array into the existing output object fbn at the end of runModel2

       -,- a,- A,- -,b -,B a,b A,B A,b a,B
SS1SS2   0   0   0   0   0   0   0   0   0
SS1RS2  NA  NA  NA  NA  NA  NA  NA  NA  NA
SS1RR2  NA  NA  NA  NA  NA  NA  NA  NA  NA
RS1SS2  NA  NA  NA  NA  NA  NA  NA  NA  NA
RS1RS2  NA  NA  NA  NA  NA  NA  NA  NA  NA
RS1RR2  NA  NA  NA  NA  NA  NA  NA  NA  NA
RR1SS2  NA  NA  NA  NA  NA  NA  NA  NA  NA
RR1RS2  NA  NA  NA  NA  NA  NA  NA  NA  NA
RR1RR2  NA  NA  NA  NA  NA  NA  NA  NA  NA

STill not quite working ...

**1/5/15** 9-14.30 5.5hrs 3.15-6 2.75 8.25hrs

Fitness matrix different in my refactored version. But do I know which is correct ?

Browse[2]> fbn
       -,- a,- A,- -,b -,B a,b A,B A,b a,B
SS1SS2   1   0   0   0   0   0   0   0   0
SS1RS2   1   0   0   0   0   0   0   0   0
SS1RR2   1   0   0   0   0   0   0   0   0
RS1SS2   1   0   0   0   0   0   0   0   0
RS1RS2   1   0   0   0   0   0   1   0   0
RS1RR2   1   0   0   0   0   0   1   0   0
RR1SS2   1   0   0   0   0   0   0   0   0
RR1RS2   1   0   0   0   0   0   1   0   0
RR1RR2   1   0   0   0   0   0   1   0   0
Browse[2]> fbn2
       -,- a,- A,- -,b -,B a,b A,B A,b a,B
SS1SS2   0   0   0   0   0   0   0   0   0
SS1RS2   0   0   0   0   0   0   0   0   0
SS1RR2   0   0   0   0   0   0   0   0   0
RS1SS2   0   0   0   0   0   0   0   0   0
RS1RS2   0   0   0   0   0   0   0   0   0
RS1RR2   0   0   0   0   0   0   0   0   0
RR1SS2   0   0   0   0   0   0   0   0   0
RR1RS2   0   0   0   0   0   0   0   0   0
RR1RR2   0   0   0   0   0   0   0   0   0

Aha difference is because Wniche just has 0s in, may be a problem earlier in the program.
Aha issue is that Wloci is not read in yet.

Reading in of Wloci can also be refactored, but I first need to read in h,z,phi & s as arrays.

Then I can try to refactor this :
```
    ## Calculated fitnesses ####

    # absence of insecticide
    ## fitness of SS in absence of insecticide is entered above as a parameter
    W.RS1_00 <- 1 - (h.RS1_00 * z.RR1_00)
    W.RR1_00 <- 1 - z.RR1_00

    W.RS2_00 <- 1 - (h.RS2_00 * z.RR2_00)
    W.RR2_00 <- 1 - z.RR2_00

    # low levels of insecticide a
    W.SS1_a0 <- 1 - phi.SS1_a0
    W.RS1_a0 <- W.SS1_a0 + (h.RS1_a0 * s.RR1_a0)
    W.RR1_a0 <- W.SS1_a0 + s.RR1_a0

    # high levels of insecticide A
    W.SS1_A0 <- 1 - phi.SS1_A0
    W.RS1_A0 <- W.SS1_A0 + (h.RS1_A0 * s.RR1_A0)
    W.RR1_A0 <- W.SS1_A0 + s.RR1_A0

    # low levels of insecticide b
    W.SS2_0b <- 1 - phi.SS2_0b
    W.RS2_0b <- W.SS2_0b + (h.RS2_0b * s.RR2_0b)
    W.RR2_0b <- W.SS2_0b + s.RR2_0b

    # high levels of insecticide B
    W.SS2_0B <- 1 - phi.SS2_0B
    W.RS2_0B <- W.SS2_0B + (h.RS2_0B * s.RR2_0B)
    W.RR2_0B <- W.SS2_0B + s.RR2_0B
```

h only has 6 values 2*3. Because it's only for heterozygous loci & niches.

```
    # h = dominance coefficient
    h.RS1_00 <- input[32,i]
    h.RS1_a0 <- input[33,i]
    h.RS1_A0 <- input[34,i]

    h.RS2_00 <- input[35,i]
    h.RS2_0b <- input[36,i]
    h.RS2_0B <- input[37,i]
```

Perhaps best to store as
h[locusNum,niche1,niche2]

This nearly works, but includes un-needed homozygous niches aa etc 
createArray( loci=c('RS1','RS2'), niche1=c('0','a','A'), niche2=c('0','b','B') )

I could add a niches arg to createArray()
createArray( loci=c('RS1','RS2'), niches=c('00','a0','A0') )

BUT no even that doesn't work. 
Because it only needs to contain the insecticide niche associated with the locus.

Actually it just needs :
h[locusNum, nicheLevel] : where insecticide is no, lo, hi
... seeing how that works.

and what about z - fitness cost of resistance allele in no insecticide
z only has 2 values
    z.RR1_00 <- input[42,i]
    z.RR2_00 <- input[43,i]
z[locusNum]
Because by definition it is the fitness cost of one allele in absence of the corresponding insecticide. (so `RR1_00` is a little misleading `z.R1_0` more like it)

phi - fitness of SS in each insecticide/concentration 
just 4 values
    phi.SS1_a0 <- input[26,i]
    phi.SS1_A0 <- input[27,i]

    phi.SS2_0b <- input[28,i]
    phi.SS2_0B <- input[29,i]

phi[locusNum, nicheLevel] where nicheLevel is lo, hi

I have Wloci being 54
> str(Wloci)
 num [1:6, 1:3, 1:3] 0 0 0 0 0 0 0 0 0 0 ...
 - attr(*, "dimnames")=List of 3
  ..$ loci  : chr [1:6] "SS1" "RS1" "RR1" "SS2" ...
  ..$ niche1: chr [1:3] "0" "a" "A"
  ..$ niche2: chr [1:3] "0" "b" "B"
> length(Wloci)
[1] 54

but actually I think it only needs 1 niche that corresponds to the chosen locus.
for locus1 it only has as and locus2 it only has bs

look at how Wloci is used to think about how best to structure

This is where it is used 
`Wniche[locus1,locus2,niche1,niche2] <- Wloci[locus1,niche1,niche2] * Wloci[locus2,niche1,niche2]`
I think I need to change from 
`Wloci[locus1,niche1,niche2] * Wloci[locus2,niche1,niche2]`
to
`Wloci[locus1,niche1] * Wloci[locus2,niche2]`

I might want to change niche names to no,lo,hi throughout. Having them different for each locus (0aA & 0bB) causes the code difficulty.

But that is a bigger change making it more difficult to check that the code logic remains the same.

A temporray solution could be to have a lookup from nicheLevel to niche.
e.g. no,lo,hi to 0,a,A & 0,b,B for locus1 & 2 respectively

Later maybe
Wloci[(locusNum),allele=(SS,SR,RR),exposure=(no,lo,hi)]
for now
Wloci[loci=(SS1,SR1,RR1,SS2,SR2,RR2),exposure=(no,lo,hi)]

s selection coeeficient : just 4 values
    # s = selection coefficient
    s.RR1_a0 <- input[38,i]
    s.RR1_A0 <- input[39,i]

    s.RR2_0b <- input[40,i]
    s.RR2_0B <- input[41,i]

It's the selection coefficient of resistance in presence of the corresponding insecticide (at lo or hi)
s[locusNum,exposure]
but exposure is only lo or hi. Would a -ve value here for no be the same as z (the fitness cost of resistance in absence of insecticide ?)

cool I think I have generic createArray2 function working

I'm getting closer to fbn & fbn2 being the same

Differences in fbn2:
1) The 00 column has 0s, in fbn it is all 1s
2) Column aB has the values that should be in AB

e.g. SS1RR2
W.SS1RR2_00
Wniche['SS1','RR2','0','0']

Wloci[locus2,exposure2] == 0
but
W.RR2_00 == 1

aha think due to typo here :
Wloci[ paste0('RR',locusNum), 'no'] <- 1 - z[locusNum]

hurrah fixed first problem, now on to 2)

Wniche looks correct
So i think it may be due to ordering of columns in fbn
Which goes AB, Ab, aB, where I think I assume aB, Ab, AB

Wniche['RR1','RR2',,]
      niche2
niche1 0 b B
     0 1 0 0
     a 0 0 0
     A 0 0 1

as.vector( Wniche['RR1','RR2',,])
[1] 1 0 0 0 0 0 0 0 1

It may be tricky for me to get output in exactly the same order as Beths.
How can I use the niche names to name the output matrix ?

Difference in orders
colnames(fbn) <- c("-,-", "a,-", "A,-", "-,b", "-,B", "a,b", "A,B", "A,b", "a,B")
my order is         00     a0     A0     0b     ab     Ab     0B    aB     AB

which is 1,2,3,4,6,8,5,9,7

I might just need to allow that the fitness list is different in my refactor test ?

Still how can I name the output columns from the data ?
Fixed !!

Then I started to delete some of the old code, and found that it broke some things I didn't want to break yet.

Particularly this which suggests I need to get into the cis/trans issue soon:

W.bar.m <- Windiv['m']
W.bar.m <- (f.m.SS1SS2 * W.m.SS1SS2) ...

revert changes for now and come back to

@@

1. what to call createArray2() ? maybe namedArray, or arrayNamed() ? could even put back in the option to pass the fill value ? submit to stackoverflow ?

@@

Different W fitness variables
`W.RR1_00 : [locus, niche1, niche2]` single locus (Wloci)
`W.RR1SS2_0b :   [locus1, locus2, niche1, niche2]` niche (Wniche)
`W.m.SS1SS2 : [sex, locus1, locus2]` sex (Windiv)

### refactoring, what I have done
Replaced the following collections of individual variables with arrays :
(To reduce code volume, reduce risk of mistyped variable names and allow objects to be passed between functions. Also potentially allows it to cope with more loci and/or niches later).

```
a.m_a0 ... *18 2*3*3  to: a[sex, niche1, niche2]
W.m.SS1SS2 ... *18 2*3*3 to:  Windiv[sex, locus1, locus2]
W.RR1SS2_0b ... *81 3*3*3*3 to: Wniche[locus1, locus2, niche1, niche2]

still todo
W.RR1_00 etc. with Wloci[loci, niche1, niche2]

because of cis/trans:
f.m.SS1SS2 ... *19 2*3*3+1 to: f[sex, genotype]

```
Want to do:
1. rename a to exposure
1. rename W to ?fitness : fitnessLocus, fitnessNiche and fitnessIndiv
